Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-02-26 parallel-tractability-J-v9/response.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{url}

\let\quoteOld\quote
\let\endquoteOld\endquote
\renewenvironment{quote}{\quoteOld\itshape}{\endquoteOld}

% use the following command line to convert to standard text file
% pandoc reviews.tex -o rebuttal.txt
\setlength{\parindent}{0pt}

\usepackage[usenames, dvipsnames]{color}
\newcommand{\BG}[1]{\textcolor{red}{BG: #1}}
%\renewcommand{\BG}[1]{}

@@

\begin{document}

We thank the reviewers for the time and effort invested in the
detailed reviews and the valuable suggestions and comments to improve
the quality of the paper. The paper has been carefully proof-read and,
hopefully, all wording and structural issues are fixed now. We also
fixed all mentioned minor issues. Below we reply to the other points
raised by the reviewers for which we completely revised the evaluation
section.

@@

\section{Rebuttal to Review 1}

\begin{quote}
There should, however, be some discussion somewhere in the paper on the
relationship between path length and KB size.   Even if a KB uses a
non-parallelizable ontology language this doesn't matter much if the paths
that cause problems are short compared to the size of the KB.  The
non-parallel aspects of path processing of short paths will not be noticable
with a small number of threads (certainly up to hundreds) and maybe not even
for very many threads (because of general overhead).
\end{quote}

We added a parallel complexity analysis for Algorithm~$\mathsf{A}_{\text{bsc}}$
and Algorithm~$\mathsf{A}_{\text{opt}}$
in Section~3.3 and Section~3.4 respectively.
Based on the analysis and Theorem~1, we have that the parallel complexity
of Algorithm~$\mathsf{A}_{\text{bsc}}$ depends only on the graph depth,
while the parallel complexity of Algorithm~$\mathsf{A}_{\text{opt}}$
depends on both of the graph depth and the size of the input ontology.\\

We added an evaluation and analysis of the role of the graph depth in Section~6.3.
The experimental results show that, when the graph depth is comparable to
the size of the given ontology, it determines the reasoning time.

@@

\begin{quote}
There is no definition of what the parents of a node in a graph are.
However, it is quite obvious what the definition should be.
\end{quote}

We added the definition of the parents of a node in a materialization graph in Definition~2.

\begin{quote}
Example 6 does need a bit of revision.  It is not the case that Oex3 cannot
be materialzed tractably in parallel.  Instead there has to be a class of
ontologies that cannot be so materialized.
\end{quote}

We revised Example~6 (see the paragraph following Example~6) and also
and Example~1 (see the last paragraph of Section~3.2).

@@

\begin{quote}
It would be much better if the ontology names corresponded to the example
numbers, i.e., the ontology in Example 6 was Oex6.
\end{quote}

We made the ontology subscripts correspond to the example numbers
for Example~5, Example~6, Example~7 and Example~9.

\begin{quote}
The notion of a simple concept is very much tied to the algorithm developed
in the paper.  How often do actual ontologies have constructs that could
make them intractable that are rended tractable by having simple concepts?
If there are none, then this class is uninteresting, particularly as the
paper makes such strong claims about worrying about particular ontologies.
This is different from an ontology only containing simple concepts.
\end{quote}

We investigated many real ontologies from different data source. These ontologies cover
several domains and only contain simple concepts (see
Section~6.1). \BG{Sec.~6.1 now only covers LUBM and YAGO. I don't think the
current answer is convincing. Did this refer to the old Sec.~6.1 with
the many more ontologies?}\\

In this work, we focused on exploring the properties
that make ontologies tractable in parallel. Thus, we did not study how to make ontologies
tractable in parallel by having simple concepts or making concepts
simple. This could be a direction for future work.

@@

\begin{quote}
The paper does a bad job of describing why materialization in RDFS is not
parallizable.  RDFS KBs have an infinite number of consequences, so it is
certainly not possible to materialize all of them quickly.  There are,
however, well known tricks that can be used to generate a finite
representation of this infinite set of consequences.  However, even this
cannot be done quickly in parallel.  In fact, even RDFS entailment cannot be
done quickly in parallel, because RDFS allows for a kind of twisted
interaction (which I think it similar to twisted paths in the paper) when
computing subproperty relationships.  Of course, having subproperties of
rdfs:subPropertyOf is not a usual thing to do in RDFS.  The paper needs to
be much more clear on its relationship to RDFS.
\end{quote}

We gave an example that RDFS entailment may lead to the situation of path twisting in Section~4.4 of the revision.
We also discussed in what cases RDFS ontologies are tractable in parallel.

\begin{quote}
The paper mentions YAGO early on but does not indicate why YAGO admits good
performance for parallel reasoning.  YAGO uses a very simple ontology
language - RDFS plus two simple extensions.  It is thus a prime candidate
for close examination to show how practical results differ from worst-case
theoretical ones.  However, the paper does not have any examination of
reasoning on YAGO.  Such an examination is very much called for.  Similarly
an examination of the performance of other large RDFS-based KBs, such as
DBpedia, would be very useful.   If YAGO is not benchmarked then it needs to
be dropped entirely.   Similarly for LUBM.
\end{quote}

We extended Section~6.1 to further discuss in which cases
the two ontologies, YAGO and LUBM, are tractable in parallel.
We added an evaluation for these ontologies in Section~6.2.

@@

\begin{quote}
The benchmarking part of the paper is very limited and extremely hard to
follow.    Only eight ontologies were benchmarked.  Why were these eight
chosen?  Only considering at most eight threads is much too constricting.
\end{quote}

The eight ontologies were selected since they satisfy the properties for parallel tractability.
We used them to examine the effects of parallel tractability.\\

In the revision, we did not continue to use the eight ontologies.
We modified the LUBM ontology to generate ontologies of different graph depths.
The experimental results were clearer for explaining the issues of graph depths
and parallel tractability.\\

We used up to 24 threads in the revised experiments.

@@

\begin{quote}
More detail is needed on how the KBs were constructed.  Chain length is a
very important aspect of the KBs, but nothing is said of how the instances
in the KB are connected together.  It is not stated which ontologies belong
to which DL.
\end{quote}

In the original experiments, we did not
construct these eight ontologies for a certain purpose, e.g., to
make the graph depth deeper. The main purpose was to closely reflect real-world scenarios. \\

In the revision, we modified the LUBM benchmark to generate ontologies of
different graph depths. We further provide a description of how these ontologies
were constructed.

@@

\begin{quote}
Figure 5 is very misleading because it uses different scales for the two
systems.   There are several immediate questions that arise from this Figure
that are poorly handled in the paper.  First, why does going from one thread
to two make such a difference for ParallelDHL?  Without a convincing
explanation for this surprising aspect of the benchmarking it is hard to
trust any of the rest of the benchmarking.  The other surprising aspect of
Figure 5 is that RDFox speeds up so little.  There is a short explanation of
why this is so, but I would have liked to see a longer one.
\end{quote}

In the original experiments, ParallelDHL required a large amount of time for computing
the relation $S_{\textit{\!\tiny rch}}$ when only one thread is being allocated.
With several threads allocated, ParallelDHL has a better efficiency.
This also resulted in the difference of the speedups between RDFox and ParallelDHL.

@@

\begin{quote}
The computation of the speedup numbers is wholely unexplained.  Why should
this number have any importance?
\end{quote}

We used the speedup numbers to show different performances between the
ontologies in $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$ and
that not in $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$.\\

In the revision, we provide the formula for the speedup and analyze
the speedup numbers.

@@

\begin{quote}
There is a claim that the benchmarking distinguishes between the Ddhlo and
D-dhlo.  However, Figure 5 doesn't really show this well, if at all.   For
example, why should processing slow down with more threads at all?  Further,
it appears that even for several of the more well-behaved ontologies that there
is very little speedup between 6 and 8 processors.
\end{quote}

\BG{The answer is not very much to the point and just a repetition of
  a previous answer.}
In order to make the original experiments close to real-world scenarios, we did not
construct the eight ontologies for some special purpose, e.g., to
make the graph depth deeper.
The differences between the ontologies in
$\mathcal{D}_{\textit{\text{dhl}}(\circ)}$ and those not in
$\mathcal{D}_{\textit{\text{dhl}}(\circ)}$ were not clear
for the test ontologies.\\

In the revision, we did not continue to use the eight ontologies.
We modified the LUBM ontology to generate ontologies of different graph depths.
The experimental results were clearer for explaining the issues of graph depths
and parallel tractability.

@@

\begin{quote}
The benchmarking should be redone with more ontologies and with more
threads.  As well, there needs to be better analysis of where parallelism is
failing.  For example, if chains are short, then there should be no problem
in achieving near-perfect thread utilization, ignoring memory contention
problems.
\end{quote}

We added a detailed analysis for LUBM and YAGO and the evaluation of
graph depths in the revision.

@@

Indeed, the abstract has been shortened significantly now.

\section{Rebuttal to Review 2}

\begin{quote}
First, in the discussion under Lemma 1, it says $|P^*|$ is polynomial in the size of $P$, which is imprecise, as $|P^*|$ is polynomial in the size of $I$ but not necessarily in the size of $R$. This is not a big issue if one considers NC for data complexity. What is less straightforward and more of concern is the statement that Step~2 costs only one time unit. As Step~2 involves checking the applicability of $B_1, \ldots, B_n \to H$ in $G$ and updating $G$, it depends on the size of $G$ and the size of $G$ is not a constant w.r.t.\ that of $I$. Footnote~4 suggests checking applicability of the rule can be done in constant time via an index. What about updating $G$ and maintaining the index?
\end{quote}

We rewrote the paragraph under Lemma~1 to analyze the parallel complexity of Algorithm~$\mathsf{A}_{\text{bsc}}$.
We made it clear that $|P^*|$ is polynomial in the size of $I$ for a class of datalog programs.\\

The statement that Step~2 costs only one time unit is based on the context of one processor. Since one processor
is allocated one rule instance (see the first paragraph of Algorithm~$\mathsf{A}_{\text{bsc}}$),
the procedure can be completed in constant time units. \\

The constant time of checking applicability by a processor is based on an assumption:
for any datalog program $P=\langle R, \textbf{I}\rangle$,
any substitution of some atom and any rule instance in $P^*$ can be mapped to a unique memory
location; further, a one-to-one relation can be established between processors and rule instances.
Under this assumption, a processor can check the applicability of its corresponding
rule instance and access the state of an atom occurring in this rule instance in constant time.
This assumption also applies to analyzing the computation of transitive closures \cite{Allender07} and
the problem of boolean matrix multiplication \cite{Raymond95}. Without this assumption, the program
costs at least linear time to load the inputs. Thus, the requirements of the \texttt{NC} algorithms
cannot be satisfied.\\

We provide a detailed description of the above assumption before Algorithm~$\mathsf{A}_{\text{bsc}}$.

@@

\begin{quote}
The computation of the rch relation $S_{\text{rch}}$ ($\dag$) looks problematic, as it requires ``H has been added to G''. If it were the case, then in Example~4 the $S_{\text{rch}}$ would always be empty, as none of the rule heads $H$ would be added to $G$. Again, the complexity analysis of $A_{\text{opt}}$ is unclear and possibly flawed. Step~2 uses an NC algorithm to compute the transitive closure of $S_{\text{rch}}$, which runs in poly-logarithmic time w.r.t.\ the size of $S_{\text{rch}}$ not necessarily so w.r.t.\ the size of $I$. What is the size of $S_{\text{rch}}$ w.r.t.\ that of $I$? Similarly as above, it involves updating $G$, which depends on the sizes of both $G$ and $S_{\text{rch}}$. Would it be still in poly-logarithmic time?
\end{quote}

The requirement ``H has been added to G'' in the description of computing the rch relation $S_{\text{rch}}$ ($\dag$) is indeed
incorrect. It should be ``H has \textbf{not} been added to G''. We fixed this error.\\

We rewrote the paragraph following Lemma~2 to give a more detailed
analysis of the parallel complexity of Algorithm~$\mathsf{A}_{\text{opt}}$.
In Step~2 of Algorithm~$\mathsf{A}_{\text{opt}}$, an \texttt{NC} algorithm is used to compute
the transitive closure of $S_{\text{rch}}$.
It can be checked that the scale of $S_{\text{rch}}$ is polynomial in the size of inputs.
The total computing time is still in poly-logarithmic time (see the detailed analysis in the revision).

@@

\begin{quote}
Finally, the algorithms assume each processor handling a rule instantiation. I wonder how it is implemented in ParallelDHL.
\end{quote}

We added a discussion of how each processor handles rule instantiations at the beginning of Section~6.2.

% We extended Section~6.1 to further discuss in what cases
% the two ontologies, YAGO and LUBM, are tractable in parallel.
% We also added the evaluation for these two kinds of ontologies in Section~6.2.\\

% The eight ontologies were selected since they satisfy the properties for parallel tractability.
% We used them to examine the effects of parallel tractability.
% In the revision, we did not continue to use the eight ontologies.
% We modified the LUBM ontology to generate ontologies of different graph depths.
% We gave the description of how these ontologies
% were constructed.\\

% In the new experiments, the results were clearer for explaining the issues of graph depths
% and parallel tractability.

@@

\bibliographystyle{elsarticle-num}
\bibliography{response}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2018-02-27 parallel-tractability-J-v9/appendix.tex
revised Appendix E-I

In Stage~1, the rules of the forms (R1-R3)
are applied to add new nodes to the constructed materialization graph.
We can observe that rule (R3) is used for computing the transitive
closure over roles.
As mentioned before, there exists an \texttt{NC} algorithm for
computing the transitive closure.
Inspired by this, we can prove that \emph{$\mathsf{A}_{\text{opt}}^\psi$
handles Stage~1 for some poly-logarithmical function $\psi$}.
The proof of this result can be shown by separately considering non-transitive roles and
roles that are transitive or influenced by transitive roles.
Specifically, we say that role $R$ is \emph{transitively influenced} (TI)
if (1) $R\circ R\sqsubseteq R\in\mathcal{R}$; or
(2) there exists a TI role $R'$ such that $R'\sqsubseteq_*R$ or $R'\sqsubseteq_*R^-$.
We say that a role is an NTI (non-transitively influenced) role if it
is not a TI role.
\BG{What you call (non-)transitively influenced is normally called
  (non-)simple (note also that it is unclear where your $\mathcal{R}$
  comes from, well, one has to guess from the RBox of the considered
  ontology). I suggest to stick to the standard naming conventions and
  use (also in the later parts). You have defined simple roles in
  Sec.~2.3 already. }% \\
% A role $R$ is \emph{non-simple} w.r.t.\ $\mathcal{O}$ if:
% \begin{enumerate}
% \item $\mathcal{R}$ contains an axiom $R \circ R \sqsubseteq R$,
% \item there exists a non-simple role $R'$ such that $R' \sqsubseteq_*
%   R$ or $R' \sqsubseteq_* R^-$.
% \end{enumerate}
% All other roles are \emph{simple}. 
% }
For each role $R\in\textbf{R}$, we further define a set $\delta_R$,
where \textbf{R} is the set of roles (role names and their inverses)
in the considered ontology, as follows:

@@

Let $\delta^*_R$ be the transitive closure of $\delta_R$ where $R$ is
a transitive role \BG{Really just transitive or non-simple (implicitly transitive)?}.
We then have the following lemma.

\begin{lemma}\label{lemma:ti}
  \BG{It is unclear what $P$ is. I assume the datalog translation of
    $\mathcal{O}$, which was also rather implicit, but I made that
    already more explicit. I suggest to add:\\Let
    $P=\langle R, \textbf{I}\rangle$ be the datalog program for an
    ontology
    $\mathcal{O}=\langle\mathcal{T},\mathcal{R},\mathcal{A}\rangle$. We
  then have that }
  $P\models R(a,b)$ implies: (1) if $R$ is an NTI role,
  $R(a,b)\in\delta_R$; (2) if $R$ is a transitive role,
  $R(a,b)\in\delta^*_{R}$; (3) if $R$ is a TI role, then
  $R(a,b)\in\delta_R$, or there exists a transitive role $R'$ such
  that $R'\sqsubseteq_* R$ and $R'(a,b)\in\delta^*_{R'}$.
\end{lemma}

Note that, for all roles $R$, $\delta_R$ can be computed by only applying (R1) and (R2).
In this sense, Lemma~\ref{lemma:ti} also indicates that:
(1) for each implicit node $R(a,b)$ where $R$ is an NTI role,
it can be added to a materialization graph by only applying (R1) and (R2);
(2) for each transitive role $R$, all implicit nodes of the form $R(a,b)$ are in the transitive closure $\delta^*_{R}$,
which can be computed by an \texttt{NC} algorithm on $\delta_{R}$;
(3) for each role $R$ that is a TI role but not a transitive role,
one can further perform~(R1) and~(R2) iteratively based on all transitive closures $\delta^*_{R'}$
where $R'$ is a transitive role.
Since all nodes generated by only applying~(R1) and~(R2) have SWD paths,
the computations of~(1) and~(3) can be handled by $\mathsf{A}_{\text{opt}}^\psi$.
Further, the transitive computation in part~(2) can also be handled by $\mathsf{A}_{\text{opt}}^\psi$.
Thus, there exists a poly-logarithmical function $\psi$
such that $\mathsf{A}_{\text{opt}}^\psi$ handles Stage~1.

The results of Stage~1 serve as the facts of Stage~2. In other words, all binary atoms
are the original nodes in Stage~2. This also means that the rules of form~(T1) and~(T3)
generate nodes with SWD paths. Due to the simple-concept restriction, for each rule
instantiation $A_1(a),A_2(a)\rightarrow B(a)$, either $A_1(a)$ or $A_2(a)$ always has an SWD path.
Thus, Stage~2 can be handled by $\mathsf{A}_{\text{opt}}$ in at most two iterations.
Based on the above analysis, this theorem holds.\BG{I assume that ``this
  theorem'' refers to Theorem~\ref{theorem:dhl}. I would prefer to
  make that explicit. }

@@

\emph{Base case}. If $R(a,b)\in\textbf{I}$, $R(a,b)\in\delta_{R}$. In this case,
regardless of whether $R$ is a TI or an NTI role, all of (1), (2) and (3) hold.

\emph{Inductive case}. We first analyze the case where $R$ is an NTI role.

\begin{enumerate}[leftmargin=12ex,label=Case~1.\arabic*, ref=Case~1.\arabic*]
\item $R(a,b)$ is derived by applying the rule (R1), w.l.o.g., $R'(a,b)\rightarrow R(a,b)$,
    which also means that $R'\sqsubseteq R\in\mathcal{R}$. Since $R$ is an NTI role, $R'$ has to
    be an NTI role according to Definition~\ref{def:deltaRdhl}. According to the induction hypothesis,
    $R'(a,b)\in\delta_{R'}$ (Case~1.1.1). If $R'(a,b)\in\mathcal{A}$, $R(a,b)\in\delta_{R}$
    holds according to Definition~\ref{def:deltaRdhl} (Case~1.1.2). If there exists some
    role $R''$ such that $R''(a,b)\in\mathcal{A}$ and $R''\sqsubseteq R'$, we also have
    $R(a,b)\in\delta_{R}$ since $R''\sqsubseteq_* R$ holds (Case~1.1.3). If there exists some
    role $R''$ such that $R''(b,a)\in\mathcal{A}$ and $R''\sqsubseteq R'^-$, the case is
    similar to Case~1.1.2.\label{dhl:NTIcaseR1}

@@

\item $R(a,b)$ is derived by applying the rule (R3). This case is impossible, since $R$ is
    an NTI role.
\end{enumerate}

@@

\begin{enumerate}[leftmargin=12ex,label=Case~2.\arabic*, ref=Case~2.\arabic*]
\item $R(a,b)$ is derived by applying the rule (R1), w.l.o.g., $R'(a,b)\rightarrow R(a,b)$,
    which also means that $R'\sqsubseteq R\in\mathcal{R}$.\label{dhl:TcaseR1}
    \begin{enumerate}[leftmargin=8ex,label=Case~2.1.\arabic*]
    \item $R'$ is an NTI role. By the induction hypothesis, $R'(a,b)\in\delta_{R'}$.
        Similar to Case~1.1.1, $R(a,b)\in\delta_{R}$ holds.
        Obviously, $R(a,b)\in\delta_{R}^*$ also holds.

    \item $R'$ is a TI role. By the induction hypothesis, if $R'(a,b)\in\delta_{R'}$,
        then $R(a,b)\in\delta_R$ also holds. On the other hand, if there exists a transitive role
        $R''$ (if $R'$ is a transitive role, then $R''\equiv R'$) such that
        $R''\sqsubseteq_* R'$ and $R''(a,b)\in\delta_{R''}^*$. Since $\delta_{R''}^*$ is
        the transitive closure of $\delta_{R''}$, we then have that there must
        exist such atoms ($R''(a,c_1),R''(c_1,c_2),...,R''(c_n,b)$) in $\delta_{R''}$.
        Further, due to $R''\sqsubseteq_* R$ and Definition~\ref{def:deltaRdhl},
        we have $R(a,c_1),R(c_1,c_2),...,$ $R(c_n,b)\in\delta_{R}$. Thus, $R(a,b)\in\delta_{R}^*$
        also holds.
    \end{enumerate}

\item $R(a,b)$ is derived by applying the rule (R2), w.l.o.g., $R'(b,a)\rightarrow R(a,b)$,
    which also means $R'\sqsubseteq R^-\in\mathcal{R}$.\label{dhl:TcaseR2}
    \begin{enumerate}[leftmargin=8ex,label=Case~2.2.\arabic*,ref=Case~2.2.\arabic*]
    \item $R'$ is an NTI role. By induction hypothesis, $R'(b,a)\in\delta_{R'}$.
        Similar to \ref{dhl:NTIcaseR2}, $R(a,b)\in\delta_{R}$ holds.
        Obviously, $R(a,b)\in\delta_{R}^*$ also holds.

    \item $R'$ is a TI role. By the induction hypothesis, if $R'(b,a)\in\delta_{R'}$,
        then $R(a,b)\in\delta_R$ also holds. On the other hand, if there exists a transitive role
        $R''$ (if $R'$ is a transitive role, then $R''\equiv R'$) such that
        $R''\sqsubseteq_* R'$ and $R''(b,a)\in\delta_{R''}^*$. Since $\delta_{R''}^*$ is
        the transitive closure of $\delta_{R''}$, we then have that there must
        exist such atoms ($R''(b,c_1),R''(c_1,c_2),...,R''(c_n,a)$) in $\delta_{R''}$.
        Further, due to $R''\sqsubseteq_* R^-$ and Definition~\ref{def:deltaRdhl},
        we have $R(a,c_n),R(c_n,c_{n-1}),...,$ $R(c_1,b)\in\delta_{R}$. Thus, $R(a,b)\in\delta_{R}^*$
        also holds.
    \end{enumerate}

@@

We finally study the case where $R$ is a TI but not a transitive role.

\begin{enumerate}[leftmargin=12ex,label=Case~3.\arabic*, ref=Case~3.\arabic*]
\item $R(a,b)$ is derived by applying the rule (R1), w.l.o.g., $R'(a,b)\rightarrow R(a,b)$,
    which also means that $R'\sqsubseteq R\in\mathcal{R}$.\label{dhl:NTcaseR1}
    \begin{enumerate}[leftmargin=8ex,label=Case~3.1.\arabic*]
    \item $R'$ is an NTI role. By the induction hypothesis, $R'(a,b)\in\delta_{R'}$ holds.
        Similar to Case~1.1.1, $R(a,b)\in\delta_{R}$ holds.
        Obviously, $R(a,b)\in\delta_{R}^*$ also holds.

    \item $R'$ is a TI role. By the induction hypothesis, if $R'(a,b)\in\delta_{R'}$,
        then $R(a,b)\in\delta_R$ also holds. On the other hand, if there exists a transitive role
        $R''$ (if $R'$ is a transitive role, then $R''\equiv R'$) such that
        $R''\sqsubseteq_* R'$ and $R''(a,b)\in\delta_{R''}^*$. Since $R''\sqsubseteq_* R$ and
        $R''$ is the transitive role, third consequence in this lemma is satisfied.
    \end{enumerate}

@@

\noindent\emph{Proof}: The proof idea of this theorem is similar to that of Theorem~\ref{theorem:dhl}.
That is, we separate the materialization of DHL$(\circ)$ ontologies into two stages: in Stage~1,
all the rules of the forms (R1-R4) are exhaustively applied; in Stage~2, the rules of the forms (T1-T2)
are then applied while the results of Stage~1 serve as facts. Stage~2 is as same as that of DHL.
Thus we only consider Stage~1 here.

Our target is to show that $\mathsf{A}_{\text{opt}}^\psi$ handles Stage~1. To this end, we also
distinguish all roles by whether they are transitively influenced. Since we have to consider
complex RIAs, we re-define TI and NTI roles as follows:
a role $R$ is \emph{transitively influenced} (TI)
if (1) $R\circ R\sqsubseteq R\in\mathcal{R}$; or
(2) there exists a TI role $R'$ such that $R'\sqsubseteq_*R$ or $R'\sqsubseteq_*R^-$;
or (3) there exist a TI role $R'$ and an axiom of the form either
$R'\circ R''\sqsubseteq R$ or $R''\circ R'\sqsubseteq R$.
We say that a role is an NTI role if it is not transitively influenced.
The set $\delta_R$ for each role $R\in\textbf{R}$ is re-defined based
on a \emph{role sequence set}
that is defined as follows:

>>>>

In Stage~1, the rules of the forms (R1-R3)
are applied to add new nodes to the constructed materialization graph.
We can observe that rule (R3) is used for computing the transitive
closure over roles.
As mentioned before, there exists an \texttt{NC} algorithm for
computing the transitive closure.
Inspired by this, we can prove that \emph{$\mathsf{A}_{\text{opt}}^\psi$
handles Stage~1 for some poly-logarithmical function $\psi$}.
The proof of this results can be shown by distinguishing simple roles
and non-simple roles.
%
For each role $R\in\textbf{R}$, we further define a set $\delta_R$,
where \textbf{R} is the set of roles (role names and their inverses)
in the considered ontology, as follows:

@@

Let $\delta^*_R$ be the transitive closure of $\delta_R$ where $R$ is a transitive role.\footnote{Note that,
in DHL, a transitive role is a non-simple role, while a non-simple role may not be a transitive role.}
We then have the following lemma.

\begin{lemma}\label{lemma:ti}
  Let
    $P=\langle R, \textbf{I}\rangle$ be the datalog program for an
    ontology
    $\mathcal{O}=\langle\mathcal{T},\mathcal{R},\mathcal{A}\rangle$. We
  then have that
  $P\models R(a,b)$ implies: (1) if $R$ is a simple role,
  $R(a,b)\in\delta_R$; (2) if $R$ is a transitive role,
  $R(a,b)\in\delta^*_{R}$; (3) if $R$ is a non-simple role but not a transitive role, then
  $R(a,b)\in\delta_R$, or there exists a transitive role $R'$ such
  that $R'\sqsubseteq_* R$ and $R'(a,b)\in\delta^*_{R'}$.
\end{lemma}

Note that, for all roles $R$, $\delta_R$ can be computed by only applying (R1) and (R2).
In this sense, Lemma~\ref{lemma:ti} also indicates that:
(1) for each implicit node $R(a,b)$ where $R$ is a simple role,
it can be added to a materialization graph by only applying (R1) and (R2);
(2) for each transitive role $R$, all implicit nodes of the form $R(a,b)$ are in the transitive closure $\delta^*_{R}$,
which can be computed by an \texttt{NC} algorithm on $\delta_{R}$;
(3) for each role $R$ that is a non-simple role but not a transitive role,
one can further perform~(R1) and~(R2) iteratively based on all transitive closures $\delta^*_{R'}$
where $R'$ is a transitive role.
Since all nodes generated by only applying~(R1) and~(R2) have SWD paths,
the computations of~(1) and~(3) can be handled by $\mathsf{A}_{\text{opt}}^\psi$.
Further, the transitive computation in part~(2) can also be handled by $\mathsf{A}_{\text{opt}}^\psi$.
Thus, there exists a poly-logarithmical function $\psi$
such that $\mathsf{A}_{\text{opt}}^\psi$ handles Stage~1.

The results of Stage~1 serve as the facts of Stage~2. In other words, all binary atoms
are the original nodes in Stage~2. This also means that the rules of form~(T1) and~(T3)
generate nodes with SWD paths. Due to the simple-concept restriction, for each rule
instantiation $A_1(a),A_2(a)\rightarrow B(a)$, either $A_1(a)$ or $A_2(a)$ always has an SWD path.
Thus, Stage~2 can be handled by $\mathsf{A}_{\text{opt}}$ in at most two iterations (see Theorem~\ref{theorem:dhl}).
Based on the above analysis, this theorem holds.

@@

\emph{Base case}. If $R(a,b)\in\textbf{I}$, $R(a,b)\in\delta_{R}$ holds. In this case,
regardless of whether $R$ is a non-simple or a simple role, all of (1), (2) and (3) in Lemma~\ref{lemma:ti} hold.

\emph{Inductive case}. We first analyze the case where $R$ is a simple role.

\begin{enumerate}[leftmargin=12ex,label=Case~1.\arabic*, ref=Case~1.\arabic*]
\item $R(a,b)$ is derived by applying the rule (R1), w.l.o.g., $R'(a,b)\rightarrow R(a,b)$,
    which also means that $R'\sqsubseteq R\in\mathcal{R}$. Since $R$ is a simple role, $R'$ has to
    be a simple role according to Definition~\ref{def:deltaRdhl}. According to the induction hypothesis,
    $R'(a,b)\in\delta_{R'}$ (Case~1.1.1). If $R'(a,b)\in\mathcal{A}$, $R(a,b)\in\delta_{R}$
    holds according to Definition~\ref{def:deltaRdhl} (Case~1.1.2). If there exists some
    role $R''$ such that $R''(a,b)\in\mathcal{A}$ and $R''\sqsubseteq R'$, we also have
    $R(a,b)\in\delta_{R}$ since $R''\sqsubseteq_* R$ holds (Case~1.1.3). If there exists some
    role $R''$ such that $R''(b,a)\in\mathcal{A}$ and $R''\sqsubseteq R'^-$, the case is
    similar to Case~1.1.2.\label{dhl:NTIcaseR1}

@@

\item $R(a,b)$ is derived by applying the rule (R3). This case is impossible, since $R$ is
    a simple role.
\end{enumerate}

@@

\begin{enumerate}[leftmargin=12ex,label=Case~2.\arabic*, ref=Case~2.\arabic*]
\item $R(a,b)$ is derived by applying the rule (R1), w.l.o.g., $R'(a,b)\rightarrow R(a,b)$,
    which also means that $R'\sqsubseteq R\in\mathcal{R}$.\label{dhl:TcaseR1}
    \begin{enumerate}[leftmargin=8ex,label=Case~2.1.\arabic*]
    \item $R'$ is a simple role. By the induction hypothesis, $R'(a,b)\in\delta_{R'}$ holds.
        Similarly to Case~1.1.1, $R(a,b)\in\delta_{R}$ holds.
        Obviously, $R(a,b)\in\delta_{R}^*$ also holds.

    \item $R'$ is a non-simple role. By the induction hypothesis, if $R'(a,b)\in\delta_{R'}$ holds,
        then $R(a,b)\in\delta_R$ also holds. On the other hand, if there exists a transitive role
        $R''$ (if $R'$ is a transitive role, then $R''\equiv R'$) such that
        $R''\sqsubseteq_* R'$ and $R''(a,b)\in\delta_{R''}^*$. Since $\delta_{R''}^*$ is
        the transitive closure of $\delta_{R''}$, we then have that there must
        exist such atoms ($R''(a,c_1),R''(c_1,c_2),...,R''(c_n,b)$) in $\delta_{R''}$.
        Further, due to $R''\sqsubseteq_* R$ and Definition~\ref{def:deltaRdhl},
        we have that $R(a,c_1),R(c_1,c_2),...,$ $R(c_n,b)\in\delta_{R}$. Thus, $R(a,b)\in\delta_{R}^*$
        also holds.
    \end{enumerate}

\item $R(a,b)$ is derived by applying the rule (R2), w.l.o.g., $R'(b,a)\rightarrow R(a,b)$,
    which also means $R'\sqsubseteq R^-\in\mathcal{R}$.\label{dhl:TcaseR2}
    \begin{enumerate}[leftmargin=8ex,label=Case~2.2.\arabic*,ref=Case~2.2.\arabic*]
    \item $R'$ is a simple role. By induction hypothesis, $R'(b,a)\in\delta_{R'}$.
        Similarly to \ref{dhl:NTIcaseR2}, $R(a,b)\in\delta_{R}$ holds.
        Obviously, $R(a,b)\in\delta_{R}^*$ also holds.

    \item $R'$ is a non-simple role. By the induction hypothesis, if $R'(b,a)\in\delta_{R'}$,
        then $R(a,b)\in\delta_R$ also holds. On the other hand, if there exists a transitive role
        $R''$ (if $R'$ is a transitive role, then $R''\equiv R'$) such that
        $R''\sqsubseteq_* R'$ and $R''(b,a)\in\delta_{R''}^*$. Since $\delta_{R''}^*$ is
        the transitive closure of $\delta_{R''}$, we then have that there must
        exist such atoms ($R''(b,c_1),R''(c_1,c_2),...,R''(c_n,a)$) in $\delta_{R''}$.
        Further, due to $R''\sqsubseteq_* R^-$ and Definition~\ref{def:deltaRdhl},
        we have that $R(a,c_n),R(c_n,c_{n-1}),...,$ $R(c_1,b)\in\delta_{R}$. Thus, $R(a,b)\in\delta_{R}^*$
        also holds.
    \end{enumerate}

@@

We finally study the case where $R$ is a non-simple but not a transitive role.

\begin{enumerate}[leftmargin=12ex,label=Case~3.\arabic*, ref=Case~3.\arabic*]
\item $R(a,b)$ is derived by applying the rule (R1), w.l.o.g., $R'(a,b)\rightarrow R(a,b)$,
    which also means that $R'\sqsubseteq R\in\mathcal{R}$.\label{dhl:NTcaseR1}
    \begin{enumerate}[leftmargin=8ex,label=Case~3.1.\arabic*]
    \item $R'$ is a simple role. By the induction hypothesis, $R'(a,b)\in\delta_{R'}$ holds.
        Similarly to Case~1.1.1, $R(a,b)\in\delta_{R}$ holds.
        Obviously, $R(a,b)\in\delta_{R}^*$ also holds.

    \item $R'$ is a non-simple role. By the induction hypothesis, if $R'(a,b)\in\delta_{R'}$,
        then $R(a,b)\in\delta_R$ also holds. On the other hand, if there exists a transitive role
        $R''$ (if $R'$ is a transitive role, then $R''\equiv R'$) such that
        $R''\sqsubseteq_* R'$ and $R''(a,b)\in\delta_{R''}^*$. Since $R''\sqsubseteq_* R$ and
        $R''$ is the transitive role, third consequence in this lemma is satisfied.
    \end{enumerate}

@@

\noindent\emph{Proof}: The proof idea of this theorem is similar to that of Theorem~\ref{theorem:dhl}.
That is, we separate the materialization of DHL$(\circ)$ ontologies into two stages: in Stage~1,
all the rules of the forms (R1-R4) are exhaustively applied; in Stage~2, the rules of the forms (T1-T2)
are then applied while the results of Stage~1 serve as facts. Stage~2 is as same as that of materializing
DHL. Thus we only consider Stage~1 here.

Our target is to show that $\mathsf{A}_{\text{opt}}^\psi$ handles Stage~1. To this end, we also
distinguish all roles by whether they are simple or not. For DHL$(\circ)$, we have to consider
complex RIAs as well. Thus,
the set $\delta_R$ for each role $R\in\textbf{R}$ should be re-defined.
We give the following definition of \emph{role sequence sets} which is
used to further define the set $\delta_R$.
